---
title: "GWP Calculations"
author: "Alia Al-Haj"
output: html_document
---

Below is code going step by step through the Glbal Warming Potential Calculations in Al-Haj & Fulweiler, 2020. Please note that for GWP and SFGWP mean methane flux rate not the median. For more details about the equations, please refer to the supplementary material.

Step 1: Download data from figshare: https://doi.org/10.6084/m9.figshare.9202097.v1
```{r}
require(tidyverse)
require(phyloseq)
dat <- read_csv(file = "https://ndownloader.figshare.com/files/19422035")
```


Step 2: We did not include the Yang et al., 2018 Zhangjiang River Estuary observation in our analysis. Remove this line of code.
```{r}
dat <- dat[-107,]
```


Step 3: Calculate the average, standard error, and median of mean methane emissions for each vegetation type. As stated in the paper in the second paragraph of section 5 "Global Estimates", 

"As mean flux was most often reported in the literature, we use the mean values of observed CH4 flux ranges, when available. If mean CH4 flux was not reported, we calculated mean CH4 flux based on the reported data. The mean of these data was multiplied by the most recent maximum and minimum estimated areal extent of the respective ecosystem type (Supporing Text 3 in Data S1; Duarte, 2017; Hamilton & Casey, 2016; Mcowen et al., 2017)."

Units: umol CH4 m-2 d-1

```{r}
#Subset the data by vegetation type
mangrove <- subset(dat, dat$`Veg Type` == 'Mangrove')
salt_marsh <- subset(dat, dat$`Veg Type` == 'Saltmarsh')
seagrass <- subset(dat, dat$`Veg Type` == 'Seagrass')

#Take an average of mean fluxes for each vegetation type
mean_mangrove <- mean(mangrove$Mean)
mean_salt_marsh <- mean(salt_marsh$Mean)
mean_seagrass <- mean(seagrass$Mean)

#Calculate the standard error of the fluxes for each vegetation type
se_mangrove <- (sd(mangrove$Mean))/(sqrt(109))
se_salt_marsh <- (sd(salt_marsh$Mean))/(sqrt(85))
se_seagrass <- (sd(seagrass$Mean))/(sqrt(28))

#Calculate the median of the fluxes for each vegetation type
median_mangrove <- median(mangrove$Mean)
median_salt_marsh <-median(salt_marsh$Mean)
median_seagrass <- median(seagrass$Mean)
```


Step 4: Scale data up to yearly flux (Supporting Text 3)

Units: Tmol CH4 m-2 yr-1
```{r}
#Mean
global_mangrove <- ((mean_mangrove*365)/(1e18))
global_salt_marsh <- ((mean_salt_marsh*365)/(1e18))
global_seagrass <- ((mean_seagrass*365)/(1e18))

#Standard error
global_mangrove_se <- ((se_mangrove*365)/(1e18))
global_salt_marsh_se <- ((se_salt_marsh*365)/(1e18))
global_seagrass_se <- ((se_seagrass*365)/(1e18))

#Median
global_mangrove_med <- ((median_mangrove*365)/(1e18))
global_salt_marsh_med <- ((median_salt_marsh*365)/(1e18))
global_seagrass_med <- ((median_seagrass*365)/(1e18))
```


Step 5: Scale data up to global area (Supporting text 3)

Units: Tmol CH4 yr-1

Global area (km2):
Mangrove: 137,760 - 152,361
Salt marsh: 55,000
Seagrass: 788,000 - 1,646,788
```{r}
#convert areas to m2 from km2
min_area_mangrove <- 137760*(1e6)
max_area_mangrove <- 152361*(1e6)
area_saltmarsh <- 55000*(1e6)
min_area_seagrass <- 788000*(1e6)
max_area_seagrass <- 1646788*(1e6)

#Mean
global_mangrove_max <- global_mangrove* max_area_mangrove
global_mangrove_min <- global_mangrove* min_area_mangrove
global_salt_marsh_area <- global_salt_marsh* area_saltmarsh
global_seagrass_max <- global_seagrass* max_area_seagrass
global_seagrass_min <- global_seagrass* min_area_seagrass

#Standard error
global_mangrove_max_se <- global_mangrove_se* max_area_mangrove
global_mangrove_min_se <- global_mangrove_se* min_area_mangrove
global_salt_marsh_area_se <- global_salt_marsh_se* area_saltmarsh
global_seagrass_max_se <- global_seagrass_se* max_area_seagrass
global_seagrass_min_se <- global_seagrass_se* min_area_seagrass

#median
global_mangrove_max_med <- global_mangrove_med* max_area_mangrove
global_mangrove_min_med <- global_mangrove_med* min_area_mangrove
global_salt_marsh_area_med <- global_salt_marsh_med* area_saltmarsh
global_seagrass_max_med <- global_seagrass_med* max_area_seagrass
global_seagrass_min_med <- global_seagrass_med* min_area_seagrass

#make a table of the above data
#GMFR = global CH4 flux rate (Tmol CH4-C year-1)
GMFR_Mean <- c(global_mangrove_max, global_mangrove_min, global_salt_marsh_area, global_seagrass_max, global_seagrass_min)
GMFR_SE <- c(global_mangrove_max_se, global_mangrove_min_se, global_salt_marsh_area_se, global_seagrass_max_se, global_seagrass_min_se)
GMFR_Med <- c(global_mangrove_max_med, global_mangrove_min_med, global_salt_marsh_area_med, global_seagrass_max_med, global_seagrass_min_med)
row_names <- c("mangrove max", "mangrove min", "salt marsh","seagrass max", "seagrass min")

GMFR_df <- data.frame(row_names, GMFR_Mean, GMFR_SE, GMFR_Med)
```


Step 6: Global Warming potential with Mean and median (UPDATED SUPPORTING TEXT 5 GCB CORRECTION)

GWP (g CO2-eq)= flux rate * (16.04 Tg Ch4/mol) * 32

```{r}
#Mean
Mangrove_GWP_max <- global_mangrove_max*16.04*32
Mangrove_GWP_min <- global_mangrove_min*16.04*32
Salt_marsh_GWP_max <- global_salt_marsh_area*16.04*32
Seagrass_GWP_max <- global_seagrass_max*16.04*32
Seagrass_GWP_min <- global_seagrass_min*16.04*32
#Median
Mangrove_GWP_max_med <- global_mangrove_max_med*16.04*32
Mangrove_GWP_min_med <- global_mangrove_min_med*16.04*32
Salt_marsh_GWP_max_med <- global_salt_marsh_area_med*16.04*32
Seagrass_GWP_max_med <- global_seagrass_max_med*16.04*32
Seagrass_GWP_min_med <- global_seagrass_min_med*16.04*32

GWP <- c(Mangrove_GWP_max, Mangrove_GWP_min, Salt_marsh_GWP_max, Seagrass_GWP_max, Seagrass_GWP_min)
GWP_Med <- c(Mangrove_GWP_max_med, Mangrove_GWP_min_med, Salt_marsh_GWP_max_med, Seagrass_GWP_max_med, Seagrass_GWP_min_med)

GMFR_df <- data.frame(GMFR_df, GWP, GWP_Med)
```


Step 7: Sustained flux global warming potential with mean and median (UPDATED Supporting text 6 GCB CORRECTION)

SFGWP (g CO2-eq)= flux rate * (16.04 Tg Ch4/mol) * 45

```{r}
#Mean
Mangrove_SFGWP_max <- global_mangrove_max*16.04*45
Mangrove_SFGWP_min <- global_mangrove_min*16.04*45
Salt_marsh_SFGWP_max <- global_salt_marsh_area*16.04*45
Seagrass_SFGWP_max <- global_seagrass_max*16.04*45
Seagrass_SFGWP_min <- global_seagrass_min*16.04*45

#Median
Mangrove_SFGWP_max_med <- global_mangrove_max_med*16.04*45
Mangrove_SFGWP_min_med <- global_mangrove_min_med*16.04*45
Salt_marsh_SFGWP_max_med <- global_salt_marsh_area_med*16.04*45
Seagrass_SFGWP_max_med <- global_seagrass_max_med*16.04*45
Seagrass_SFGWP_min_med <- global_seagrass_min_med*16.04*45


SFGWP <- c(Mangrove_SFGWP_max, Mangrove_SFGWP_min, Salt_marsh_SFGWP_max, Seagrass_SFGWP_max, Seagrass_SFGWP_min)
SFGWP_Med <- c(Mangrove_SFGWP_max_med, Mangrove_SFGWP_min_med, Salt_marsh_SFGWP_max_med, Seagrass_SFGWP_max_med, Seagrass_SFGWP_min_med)


GMFR_df <- data.frame(GMFR_df, SFGWP, SFGWP_Med)
GMFR_df
```


Step 8 calculate offset:
Convert Global Carbon burial rate from Tg C yr-1 to Tg CO2 yr-1
Global C burial rate (Tg CO2 yr-1) = Global C burial rate (Tg C yr-1) * 44.01 Tg CO2/ 12.01 Tg C

```{r}
Mangrove_GCB_Min <- 31.1 * 44.01/12.01
Mangrove_GCB_Max <- 34.4 * 44.01/12.01
Saltmarsh_GCB <- 11.99 * 44.01/12.01
Seagrass_GCB_Min <- 108.74*44.01/12.01 
Seagrass_GCB_Max <- 227.26*44.01/12.01
```

Calculate Mean Offset:
Offset = GWP or SGWP/global C burial rate
```{r}
Mangrove_Offset_GWP_Min <- Mangrove_GWP_min/Mangrove_GCB_Min
Mangrove_Offset_GWP_Max <- Mangrove_GWP_max/Mangrove_GCB_Max
Mangrove_Offset_SGWP_Min <-Mangrove_SFGWP_min/Mangrove_GCB_Min
Mangrove_Offset_SGWP_Max <- Mangrove_SFGWP_max/Mangrove_GCB_Max

Saltmarsh_Offset_GWP <- Salt_marsh_GWP_max/Saltmarsh_GCB
Saltmarsh_Offset_SGWP <- Salt_marsh_SFGWP_max/Saltmarsh_GCB

Seagrass_Offset_GWP <- Seagrass_GWP_max/Seagrass_GCB_Max
Seagrass_Offset_GWP <- Seagrass_GWP_min/Seagrass_GCB_Min
Seagrass_Offset_SGWP <- Seagrass_SFGWP_max/Seagrass_GCB_Max
Seagrass_Offset_SGWP <- Seagrass_SFGWP_min/Seagrass_GCB_Min
```

Calculate Median Offset:
Convert Global median Carbon burial rate from g C m-2 yr-1 to Tg CO2 yr-1
Global C burial rate (Tg CO2 yr-1) = Global C burial rate (g C m-2 yr-1) *area * 10^-12 * 44.01 Tg CO2/ 12.01 Tg C

```{r}
Mangrove_GCB_med_min <- 154 * min_area_mangrove * (10^-12) * 44.01/12.01
Mangrove_GCB_med_Max <- 154 * max_area_mangrove * (10^-12) * 44.01/12.01
saltmarsh_GCB_med <- 139.5 * area_saltmarsh* (10^-12) * 44.01/12.01
Seagrass_GCB_med_min <- 67.75 * min_area_seagrass * (10^-12) * 44.01/12.01
Seagrass_GCB_med_Max <- 67.75 * max_area_seagrass * (10^-12) * 44.01/12.01
```

Calculate Median Offset:
Offset = GWP or SGWP/global C burial rate
```{r}
Mangrove_Offset_med_GWP_Min <- Mangrove_GWP_min_med/Mangrove_GCB_med_min
Mangrove_Offset_med_GWP_Max <- Mangrove_GWP_max_med/Mangrove_GCB_med_Max
Mangrove_Offset_med_SGWP_Min <-Mangrove_SFGWP_min_med/Mangrove_GCB_med_min
Mangrove_Offset_med_SGWP_Max <- Mangrove_SFGWP_max_med/Mangrove_GCB_med_Max

Saltmarsh_Offset_med_GWP <- Salt_marsh_GWP_max_med/saltmarsh_GCB_med
Saltmarsh_Offset_med_SGWP <- Salt_marsh_SFGWP_max_med/saltmarsh_GCB_med

Seagrass_Offset_med_GWP_max <- Seagrass_GWP_max_med/Seagrass_GCB_med_Max
Seagrass_Offset_med_GWP_min <- Seagrass_GWP_min_med/Seagrass_GCB_med_min
Seagrass_Offset_med_SGWP_max <- Seagrass_SFGWP_max_med/Seagrass_GCB_med_Max
Seagrass_Offset_med_SGWP_min <- Seagrass_SFGWP_min_med/Seagrass_GCB_med_min
```